---
title: "R Notebook"
output: html_notebook
---

# Read modules
```{r}
library(dplyr)
library(ggplot2)
library(patchwork)
library(ggVennDiagram)
library(purrr)
library(waterfalls)
library(stringr)
```

```{bash}
pwd
ls
```
# mT analysis with sample data
## Read data from tsv file
```{r}
# Data DO NOT contain header
read.table(file = './ALL_ICmT_Hv3_sample_stats.tsv', sep = '\t', header = FALSE) %>% 
  as_tibble() -> df_samples

colnames(df_samples) <- c('transcript_id',
                          'tissue',
                          'stage',
                          'technology',
                          'capture_status',
                          'biological_replicate',
                          'technical_replicate')
```

## Determine number of transcripts for ea platform (multi - both ont and pacBio)
```{r}
multi_transcripts <- df_samples %>%
  select(transcript_id, technology) %>%
  filter(technology=="ont") %>%
  select(transcript_id) %>%
  intersect(., (df_samples %>%
                  select(transcript_id, technology) %>%
                  filter(technology=="pacBioSII") %>%
                  select(transcript_id)))

technology_df <- data.frame(technology = c("multi", "ont", "pacBioSII"),
                            number_of_transcripts = c(multi_transcripts %>%
                                                        nrow(),
                                                      df_samples %>%
                                                        select(transcript_id, technology) %>%
                                                        filter(technology=="ont") %>%
                                                        select(transcript_id) %>%
                                                        setdiff(., multi_transcripts) %>%
                                                        nrow(),
                                                      df_samples %>%
                                                        select(transcript_id, technology) %>%
                                                        filter(technology=="pacBioSII") %>%
                                                        select(transcript_id) %>%
                                                        setdiff(., multi_transcripts) %>%
                                                        nrow()))

# Plot the results
technology_df %>%
  mutate(percent = prop.table(number_of_transcripts)) %>% 

  ggplot(data=.,
         aes(x=reorder(technology, -percent),
             y=percent,
             fill=technology)) + 
  geom_bar(stat="identity", position = "dodge") +
  geom_text(aes(label = scales::comma(number_of_transcripts)),
            position = position_dodge(width = 1),
            vjust=-0.5,
            size = 3) +
  theme_minimal() +
  scale_fill_brewer(palette="Set2") +
  ylab('Percentage') +
  xlab('Technology') +
  theme(legend.position = "none") +
  ggtitle(label="Number of full mT transcripts aquired using ont, pacBio or both technologies.") +
  scale_y_continuous(labels = scales::percent_format(),
                     limits = c(0,1)) -> p_technology

pdf(file = "Hv3_splicedmasterTable_refined_transcripts_by_technology.pdf", width = 6, height = 4)
p_technology



##### Stacked barplot
# Plot the results
technology_df %>%
  mutate(percent = prop.table(number_of_transcripts)) %>% 

  ggplot(data=.,
         aes(x=1,#reorder(technology, -percent),
             y=percent,
             fill=factor(technology, levels=c("ont", "multi", "pacBioSII")))) + 
  geom_bar(stat="identity") +
   geom_text(aes(label = paste0(technology, " (", scales::comma(number_of_transcripts), ")")),
             position=position_stack(vjust=0.5),
             size = 4) +
  theme_minimal() +
  scale_fill_manual(values=c("#377EB8", "#4DAF4A", "#FFFF33")) +
  ylab('Percentage') +
  xlab('Technology') +
  theme(legend.position = "none",
        axis.ticks.y = element_line(colour = "black"),
        #axis.title.x = element_blank(),
        plot.title = element_text(hjust = 0.5,
                                  size=9.0),
        panel.background = element_rect(fill = NA),
        panel.grid.major.x = element_line(colour = NA),
        panel.grid.minor.x = element_line(colour = NA),
        axis.text.x = element_blank(),
        axis.text.y = element_text(colour="black",
                                  face="bold")) +
  ggtitle(label="Number of full mT transcripts aquired\nusing ont, pacBio or both technologies.") +
  scale_y_continuous(labels = scales::percent_format(),
                     limits = c(0,1)) -> p_technology2

pdf(file = "Hv3_splicedmasterTable_refined_transcripts_by_technology_stacked.pdf", width = 3, height = 4)
p_technology2

```

## Tissues - Incremental plots

### Incremental plot for pre-capture tissues of ALL STAGES
```{r}
pre_cap_all_df <- df_samples %>% 
  filter(capture_status=="pre-capture") %>%    # Only pre-capture
  select(transcript_id, tissue) %>%            
  distinct(transcript_id, tissue)              # Leave only unique tissue per transcript


# Determine tissues with highest transcripts number 
pre_cap_all_df %>% 
  group_by(tissue) %>% 
  summarise(count=n()) %>% 
  arrange(count) %>% 
  select(tissue) %>%
  as.list()

# Lowest -> Highest
# "Placenta" "TpoolA"   "CpoolA"   "WBlood"   "iPSC"     "Testis"   "Liver"    "Heart"    "Brain"   



pre_cap_all_df %>%
  group_by(transcript_id) %>%                 
  summarise(number_of_tissues=n()) %>% 
  filter(number_of_tissues==df_samples %>%     # Leave only transcripts present in all of the tissues
           distinct(tissue) %>%                # filter(number_of_tissues=max(number_of_tissues)) to
           nrow()) %>%                         # check for anomalies
  select(transcript_id) -> smallest_common          

pre_cap_all_df %>% 
  filter(tissue=="Placenta") %>%
  select(transcript_id) %>% 
  setdiff(smallest_common) -> plac

pre_cap_all_df %>% 
  filter(tissue=="TpoolA") %>%
  select(transcript_id) %>% 
  setdiff(union(smallest_common, plac)) -> tpoola

pre_cap_all_df %>% 
  filter(tissue=="CpoolA") %>%
  select(transcript_id) %>% 
  setdiff(union(smallest_common, union(plac, tpoola))) -> cpoola


pre_cap_all_df %>% 
  filter(tissue=="WBlood") %>%
  select(transcript_id) %>% 
  setdiff(union(cpoola, union(smallest_common, union(plac, tpoola)))) -> wblood


pre_cap_all_df %>% 
  filter(tissue=="iPSC") %>%
  select(transcript_id) %>% 
  setdiff(union(wblood, union(cpoola, union(smallest_common, union(plac, tpoola))))) -> ipsc


pre_cap_all_df %>% 
  filter(tissue=="Testis") %>%
  select(transcript_id) %>% 
  setdiff(union(ipsc, union(wblood, union(cpoola, union(smallest_common, union(plac, tpoola)))))) -> testis


pre_cap_all_df %>% 
  filter(tissue=="Liver") %>%
  select(transcript_id) %>% 
  setdiff(union(testis, union(ipsc, union(wblood, union(cpoola, union(smallest_common, union(plac, tpoola))))))) -> liver


pre_cap_all_df %>% 
  filter(tissue=="Heart") %>%
  select(transcript_id) %>% 
  setdiff(union(liver, union(testis, union(ipsc, union(wblood, union(cpoola, union(smallest_common, union(plac, tpoola)))))))) -> heart


pre_cap_all_df %>% 
  filter(tissue=="Brain") %>%
  select(transcript_id) %>% 
  setdiff(union(heart, union(liver, union(testis, union(ipsc, union(wblood, union(cpoola, union(smallest_common, union(plac, tpoola))))))))) -> brain


df_waterfall <- data.frame(tissue=c("common", "Placenta", "TpoolA", "CpoolA", "WBlood","iPSC","Testis","Liver","Heart", "Brain"),
                           number=c(smallest_common %>% 
                                      nrow(),
                                    plac %>% 
                                      nrow(),
                                    tpoola %>% 
                                      nrow(),
                                    cpoola %>% 
                                      nrow(),
                                    wblood %>% 
                                      nrow(),
                                    ipsc %>%
                                      nrow(),
                                    testis %>% 
                                      nrow(),
                                    liver %>% 
                                      nrow(),
                                    heart %>% 
                                      nrow(),
                                    brain %>% 
                                      nrow())
  
)

waterfall(df_waterfall,
          calc_total = TRUE,
          #rect_width = 0.4,
          draw_lines = TRUE,
          linetype = 3,
          fill_by_sign = FALSE,
          fill_colours = c("#FF7F00","#FF7F00","#FF7F00","#FF7F00","#FF7F00","#FF7F00","#FF7F00","#FF7F00","#FF7F00","#FF7F00","#FF7F00"),
          total_rect_color = "#4DAF4A",
          total_rect_text_color = "black") +
  ylab('Percentage') +
  theme_minimal() +
  ylab('Percentage') +
  theme(axis.title.x=element_blank(),
        axis.text = element_text(colour = "black", size = 10),
        axis.ticks = element_line(colour = "black"),
        panel.background = element_rect(fill = "white", colour = "grey50"),
        legend.position = "none",
        panel.grid = element_blank()
        ) +
  scale_y_continuous(labels = function(x) paste0(round(x/sum(df_waterfall$number)*100, 0), "%"),
                     breaks = seq(0, sum(df_waterfall$number), sum(df_waterfall$number)*0.25)) -> p_all_tissue_waterfall


# # Test
# pre_cap_df %>% 
#   distinct(transcript_id) %>% 
#   nrow()

pdf(file = "Hv3_splicedmasterTable_refined_waterfall_tissue_pre", width = 7, height = 4)
p_all_tissue_waterfall

```

### Pre-capture tissues of ADULT STAGE data preparation
```{r}
pre_cap_adult_df <- df_samples %>% 
  filter(capture_status=="pre-capture") %>%    # Only pre-capture
  filter(stage=="Adult") %>% 
  select(transcript_id, tissue) %>%            
  distinct(transcript_id, tissue)              # Leave only unique tissue per transcript


# Determine tissues with highest transcripts number 
pre_cap_adult_df %>% 
  group_by(tissue) %>% 
  summarise(count=n()) %>% 
  arrange(count) %>% 
  select(tissue) %>%
  as.list()

# Lowest -> Highest
# "TpoolA" "Liver"  "CpoolA" "WBlood" "Heart"  "Brain"  "Testis"


pre_cap_adult_df %>%
  group_by(transcript_id) %>%                 
  summarise(number_of_tissues=n()) %>% 
  filter(number_of_tissues==pre_cap_adult_df %>%     # Leave only transcripts present in all of the tissues
           distinct(tissue) %>%                      # filter(number_of_tissues=max(number_of_tissues)) to
           nrow()) %>%                               # check for anomalies
  select(transcript_id) -> pre_smallest_common          


pre_cap_adult_df %>% 
  filter(tissue=="TpoolA") %>%
  select(transcript_id) %>% 
  setdiff(pre_smallest_common) -> pre_tpoola

pre_cap_adult_df %>% 
  filter(tissue=="Liver") %>%
  select(transcript_id) %>% 
  setdiff(union(pre_smallest_common, pre_tpoola)) -> pre_liver


pre_cap_adult_df %>% 
  filter(tissue=="CpoolA") %>%
  select(transcript_id) %>% 
  setdiff(union(pre_liver, union(pre_smallest_common, pre_tpoola))) -> pre_cpoola


pre_cap_adult_df %>% 
  filter(tissue=="WBlood") %>%
  select(transcript_id) %>% 
  setdiff(union(pre_cpoola, union(pre_liver, union(pre_smallest_common, pre_tpoola)))) -> pre_wblood


pre_cap_adult_df %>% 
  filter(tissue=="Heart") %>%
  select(transcript_id) %>% 
  setdiff(union(pre_wblood, union(pre_cpoola, union(pre_liver, union(pre_smallest_common, pre_tpoola))))) -> pre_heart


pre_cap_adult_df %>% 
  filter(tissue=="Brain") %>%
  select(transcript_id) %>% 
  setdiff(union(pre_heart, union(pre_wblood, union(pre_cpoola, union(pre_liver, union(pre_smallest_common, pre_tpoola)))))) -> pre_brain


pre_cap_adult_df %>% 
  filter(tissue=="Testis") %>%
  select(transcript_id) %>% 
  setdiff(union(pre_brain, union(pre_heart, union(pre_wblood, union(pre_cpoola, union(pre_liver, union(pre_smallest_common, pre_tpoola))))))) -> pre_testis


pre_adult_tissues_df <- data.frame(tissue=factor(c("common", "TpoolA", "Liver", "CpoolA","WBlood","Heart","Brain","Testis"),
                                                 levels = c("Testis","Brain","Heart","WBlood", "CpoolA", "Liver", "TpoolA", "common")),
                                   number=c(pre_smallest_common %>%
                                              nrow(),
                                            pre_tpoola %>% 
                                              nrow(),
                                            pre_liver %>% 
                                              nrow(),
                                            pre_cpoola %>% 
                                              nrow(),
                                            pre_wblood %>%
                                              nrow(),
                                            pre_heart %>% 
                                              nrow(),
                                            pre_brain %>% 
                                              nrow(),
                                            pre_testis %>% 
                                              nrow()),
                                   capture_status=c("pre-capture")
)

```

### Post-capture tissues of ADULT STAGE data preparation
```{r}
post_cap_adult_df <- df_samples %>% 
  filter(capture_status=="post-capture") %>%    # Only post-capture
  filter(stage=="Adult") %>% 
  select(transcript_id, tissue) %>%            
  distinct(transcript_id, tissue)              # Leave only unique tissue per transcript


# Determine tissues with highest transcripts number 
post_cap_adult_df %>% 
  group_by(tissue) %>% 
  summarise(count=n()) %>% 
  arrange(count) %>% 
  select(tissue) %>%
  as.list()

# Lowest -> Highest
# "TpoolA" "WBlood" "Heart"  "CpoolA" "Liver"  "Brain"  "Testis"

post_cap_adult_df %>%
  group_by(transcript_id) %>%                 
  summarise(number_of_tissues=n()) %>% 
  filter(number_of_tissues==post_cap_adult_df %>%     # Leave only transcripts present in all of the tissues
           distinct(tissue) %>%                       # filter(number_of_tissues=max(number_of_tissues)) to
           nrow()) %>%                                # check for anomalies
  select(transcript_id) -> post_smallest_common          


post_cap_adult_df %>% 
  filter(tissue=="TpoolA") %>%
  select(transcript_id) %>% 
  setdiff(post_smallest_common) -> post_tpoola

post_cap_adult_df %>% 
  filter(tissue=="WBlood") %>%
  select(transcript_id) %>% 
  setdiff(union(post_smallest_common, post_tpoola)) -> post_wblood


post_cap_adult_df %>% 
  filter(tissue=="Heart") %>%
  select(transcript_id) %>% 
  setdiff(union(post_wblood, union(post_smallest_common, post_tpoola))) -> post_heart


post_cap_adult_df %>% 
  filter(tissue=="CpoolA") %>%
  select(transcript_id) %>% 
  setdiff(union(post_heart, union(post_wblood, union(post_smallest_common, post_tpoola)))) -> post_cpoola


post_cap_adult_df %>% 
  filter(tissue=="Liver") %>%
  select(transcript_id) %>% 
  setdiff(union(post_cpoola, union(post_heart, union(post_wblood, union(post_smallest_common, post_tpoola))))) -> post_liver


post_cap_adult_df %>% 
  filter(tissue=="Brain") %>%
  select(transcript_id) %>% 
  setdiff(union(post_liver, union(post_cpoola, union(post_heart, union(post_wblood, union(post_smallest_common, post_tpoola)))))) -> post_brain


post_cap_adult_df %>% 
  filter(tissue=="Testis") %>%
  select(transcript_id) %>% 
  setdiff(union(post_brain, union(post_liver, union(post_cpoola, union(post_heart, union(post_wblood, union(post_smallest_common, post_tpoola))))))) -> post_testis


post_adult_tissues_df <- data.frame(tissue=factor(c("common", "TpoolA", "WBlood", "Heart", "CpoolA", "Liver", "Brain", "Testis"),
                                                  levels=c("Testis", "Brain", "Liver", "CpoolA", "Heart", "WBlood", "TpoolA", "common")),
                                   number=c(post_smallest_common %>%
                                              nrow(),
                                            post_tpoola %>% 
                                              nrow(),
                                            post_wblood %>% 
                                              nrow(),
                                            post_heart %>% 
                                              nrow(),
                                            post_cpoola %>%
                                              nrow(),
                                            post_liver %>% 
                                              nrow(),
                                            post_brain %>% 
                                              nrow(),
                                            post_testis %>% 
                                              nrow()),
                                   capture_status=c("post-capture")
)

```

### Create the incremental plot
```{r}
ggplot() +
  geom_bar(aes(x=pre_adult_tissues_df$capture_status,
               y=pre_adult_tissues_df$number,
               fill=pre_adult_tissues_df$tissue),
           position="stack", stat="identity") +
  geom_bar(aes(x=post_adult_tissues_df$capture_status,
               y=post_adult_tissues_df$number,
               fill=post_adult_tissues_df$tissue),
           position="stack", stat="identity") +
  geom_text(aes(x=pre_adult_tissues_df$capture_status,
                y=pre_adult_tissues_df$number,
                label = (paste0(pre_adult_tissues_df$tissue, " (", scales::comma(pre_adult_tissues_df$number), ")"))),
            position = position_stack(vjust = .5),
            size = 3) +
  geom_text(aes(x=post_adult_tissues_df$capture_status,
              y=post_adult_tissues_df$number,
              label = (paste0(post_adult_tissues_df$tissue, " (", scales::comma(post_adult_tissues_df$number), ")"))),
          position = position_stack(vjust = .5),
          size = 3) +
  theme_minimal() +
  theme(axis.text = element_text(colour = "black", size = 10),
      axis.ticks = element_line(colour = "black"),
      panel.background = element_rect(fill = "white", colour = "grey50"),
      legend.position = "none",
      panel.grid = element_blank()
      ) +
  scale_fill_brewer(palette="Set2") +
  ylab('Number of transcripts') +
  xlab('Capture status') +
  theme(legend.position = "none") + 
  scale_y_continuous(labels = scales::label_number(suffix = "k", scale = 1e-3, big.mark = ",")) -> p_tissue_incremental

pdf(file = "Hv3_splicedmasterTable_refined_tissue_incremental", width = 6, height = 4)
p_tissue_incremental

```


## Tissues - Correlation matrix
### Pre-capture
```{r}
pre_cap_all_matrix_df <- df_samples %>%
  filter(capture_status=="pre-capture") %>%    # Only pre-capture
  distinct(transcript_id, tissue, stage) %>%   # Leave only unique tissue per transcript > don't include technology
  mutate(tissue=paste0(tissue," (", stage, ")")) %>% 
  select(transcript_id, tissue) %>% 
  group_by(tissue) %>% 
  summarize(transcripts = list(transcript_id)) %>% 
  cross_join(x=., y=., suffix = c("1", "2")) %>% 
  rowwise() %>% 
  mutate(intersecting = list(intersect(x=transcripts1, y=transcripts2))) %>% 
  mutate(intersecting_no = length(intersecting)) %>% 
  mutate(percentage = intersecting_no/length(transcripts1) * 100) %>% 
  mutate(transcripts1_number = length(transcripts1))


pre_cap_all_matrix_df %>% 
  ggplot(., aes(tissue1, tissue2)) +
  geom_tile(aes(fill = percentage)) +
  #geom_text(aes(label = round(value, 1))) +
  scale_fill_gradient(low = "white", high = "red")


pre_cap_all_matrix_df %>% 
  ggplot(data=.) +
  geom_tile(data = pre_cap_all_matrix_df %>% filter(percentage == 100),
            aes(x=tissue1,
                y=tissue2), fill="#A2B46A") +     # #2491AA
  geom_tile(data = pre_cap_all_matrix_df %>% filter(percentage < 100),
            aes(x=tissue1,
                y=tissue2,
                fill = percentage)) +
  geom_text(size=5,
            aes(x=tissue1,
                y=tissue2,
                label = (scales::percent(percentage/100, accuracy = 0.1L)))) +
  geom_text(size=3.2,
            aes(x=tissue1,
                y=tissue2,
                label = paste0("\n\n",
                               "(", scales::comma(intersecting_no), "/", scales::comma(transcripts1_number), ")"))) +
  scale_fill_gradient(low = "white",
                      high = "#BDD27C") +    ##29aac7
   theme(axis.title.x = element_blank(),
         axis.title.y = element_blank(),
         axis.ticks = element_line(colour = NA),
         plot.title = element_text(hjust = 0.5),
         panel.background = element_rect(fill = NA),
         panel.grid.major = element_line(colour = NA),
         axis.text.x = element_text(colour="black",
                                    face="bold"),
         axis.text.y = element_text(colour="black",
                                    face="bold")) +
  labs(title = "Full mT transcripts shared between tissues (pre-capture).", fill = "% of transcripts") -> p_pre_cap_heatmap


pdf(file="Hv3_splicedmasterTable_refined_pre_cap_heatmap.pdf", width = 16, height = 8)
p_pre_cap_heatmap

# label = round(percentage, 1))) +
# label = scales::percent(percentage/100, accuracy = 0.1L))) +
```
### Post-capture
```{r}
post_cap_all_matrix_df <- df_samples %>%
  filter(capture_status=="post-capture") %>%    # Only post-capture
  distinct(transcript_id, tissue, stage) %>%   # Leave only unique tissue per transcript > don't include technology
  mutate(tissue=paste0(tissue," (", stage, ")")) %>% 
  select(transcript_id, tissue) %>% 
  group_by(tissue) %>% 
  summarize(transcripts = list(transcript_id)) %>% 
  cross_join(x=., y=., suffix = c("1", "2")) %>% 
  rowwise() %>% 
  mutate(intersecting = list(intersect(x=transcripts1, y=transcripts2))) %>% 
  mutate(intersecting_no = length(intersecting)) %>% 
  mutate(percentage = intersecting_no/length(transcripts1) * 100) %>% 
  mutate(transcripts1_number = length(transcripts1))

post_cap_all_matrix_df %>% 
  ggplot(data=.) +
  geom_tile(data = post_cap_all_matrix_df %>% filter(percentage == 100),
            aes(x=tissue1,
                y=tissue2), fill="#55985F") +  # #30E48D
  geom_tile(data = post_cap_all_matrix_df %>% filter(percentage < 100),
            aes(x=tissue1,
                y=tissue2,
                fill = percentage)) +
  geom_text(size=5,
            aes(x=tissue1,
                y=tissue2,
                label = (scales::percent(percentage/100, accuracy = 0.1L)))) +
  geom_text(size=3.2,
            aes(x=tissue1,
                y=tissue2,
                label = paste0("\n\n",
                               "(", scales::comma(intersecting_no), "/", scales::comma(transcripts1_number), ")"))) +
  scale_fill_gradient(low = "white",
                      high = "#66B672") +  ##2AC77B
   theme(axis.title.x = element_blank(),
         axis.title.y = element_blank(),
         axis.ticks = element_line(colour = NA),
         plot.title = element_text(hjust = 0.5),
         panel.background = element_rect(fill = NA),
         panel.grid.major = element_line(colour = NA),
         axis.text.x = element_text(colour="black",
                                    face="bold"),
         axis.text.y = element_text(colour="black",
                                    face="bold")) +
  labs(title = "Full mT transcripts shared between tissues (post-capture).", fill = "% of transcripts") -> p_post_cap_heatmap


pdf(file="Hv3_splicedmasterTable_refined_post_cap_heatmap.pdf", width = 16, height = 8)
p_post_cap_heatmap

```
### Pre&Post-capture
```{r}
preANDpost_cap_all_matrix_df <- df_samples %>%
  distinct(transcript_id, tissue, stage) %>%   # Leave only unique tissue per transcript > don't include technology
  mutate(tissue=paste0(tissue," (", stage, ")")) %>% 
  select(transcript_id, tissue) %>% 
  group_by(tissue) %>% 
  summarize(transcripts = list(transcript_id)) %>% 
  cross_join(x=., y=., suffix = c("1", "2")) %>% 
  rowwise() %>% 
  mutate(intersecting = list(intersect(x=transcripts1, y=transcripts2))) %>% 
  mutate(intersecting_no = length(intersecting)) %>% 
  mutate(percentage = intersecting_no/length(transcripts1) * 100) %>% 
  mutate(transcripts1_number = length(transcripts1))

preANDpost_cap_all_matrix_df %>% 
  ggplot(data=.) +
  geom_tile(data = preANDpost_cap_all_matrix_df %>% filter(percentage == 100),
            aes(x=tissue1,
                y=tissue2), fill="#029A4B") +  ##E48430  #208a82
  geom_tile(data = preANDpost_cap_all_matrix_df %>% filter(percentage < 100),
            aes(x=tissue1,
                y=tissue2,
                fill = percentage)) +
  geom_text(size=5,
            aes(x=tissue1,
                y=tissue2,
                label = (scales::percent(percentage/100, accuracy = 0.1L)))) +
  geom_text(size=3.2,
            aes(x=tissue1,
                y=tissue2,
                label = paste0("\n\n",
                               "(", scales::comma(intersecting_no), "/", scales::comma(transcripts1_number), ")"))) +
  scale_fill_gradient(low = "white",
                      high = "#02B659") + #C8742A #2ab3a9
   theme(axis.title.x = element_blank(),
         axis.title.y = element_blank(),
         axis.ticks = element_line(colour = NA),
         plot.title = element_text(hjust = 0.5),
         panel.background = element_rect(fill = NA),
         panel.grid.major = element_line(colour = NA),
         axis.text.x = element_text(colour="black",
                                    face="bold"),
         axis.text.y = element_text(colour="black",
                                    face="bold")) +
  labs(title = "Full mT transcripts shared between tissues (pre&post-capture).", fill = "% of transcripts") -> p_preANDpost_cap_heatmap


pdf(file="Hv3_splicedmasterTable_refined_preANDpost_cap_heatmap.pdf", width = 16, height = 8)
p_preANDpost_cap_heatmap
```








