---
title: "RNA_mT_Hv3_stats"
output:
  html_notebook: default
  pdf_document: default
---

# Load necessary libraries
```{r}
library(dplyr)
library(ggplot2)
library(patchwork)
library(purrr)
library(stringr)
```

# Basic RNA mT stats
## Read data from tsv file
```{r}
# Data contains the header
read.table(file = '~/Desktop/mT/formatted_RNA_Hv3_mT.tsv', sep = '\t', header = TRUE) %>% 
  as_tibble() -> df_rna_basic
```


## Verify samples - count how many transcripts were present only in one sample - using formatted_RNA_Hv3_mT.tsv
### Histogram
```{r}
#pdf(file = "RNA_mT_sample_analysis_pre_and_post", width = 6, height = 4)
df_rna_basic %>%
  select(transcript_id, sampleN) %>% 
  distinct(transcript_id, sampleN) %>% 
  #arrange(desc(sampleN))
  filter(sampleN < 15) %>%
  ggplot(., aes(x=sampleN)) + 
  geom_histogram(binwidth=1) +
  ggtitle("Histogram for RNA mT transcripts (pre- and post-capture)")
```
### Barplot
```{r}
#pdf(file = "RNA_mT_sample_analysis_pre_and_post.pdf", width = 6, height = 4)
df_rna_basic %>%
  select(transcript_id, sampleN) %>% 
  distinct(transcript_id, sampleN) %>% 
  group_by(sampleN) %>% 
  summarise(total_count=n_distinct(transcript_id)) %>%
  filter(sampleN <= 10) %>% 
  mutate(sampleN = as.character(sampleN)) %>% 
  add_row(sampleN=paste0(11, "-", (df_rna_basic %>% 
                                     select(sampleN) %>%
                                     max())),
          total_count=(df_rna_basic %>%
                         select(transcript_id, sampleN) %>%
                         distinct(transcript_id, sampleN) %>%
                         group_by(sampleN) %>%
                         summarise(total_count=n_distinct(transcript_id)) %>%
                         filter(sampleN > 10) %>% 
                         select(total_count) %>% 
                         sum())) %>% 
  mutate(percentage = prop.table(total_count)) %>%
  ggplot(., aes(x=factor(sampleN, levels=c("1", "2", "3", "4", "5", "6", "7",
                                           "8", "9", "10", "11-49")),
                y=percentage)) + 
  geom_bar(stat="identity", fill="steelblue") +
  geom_text(aes(label=scales::comma(total_count)), vjust=-0.3, size=3.5) +
  theme_minimal() +
  #scale_x_continuous(breaks = seq(1, 20, by = 1)) +
  scale_y_continuous(labels = scales::percent_format(),
                     limits = c(0,1)) +
  ylab('Percentage') +
  xlab('sampleN') +
  theme(legend.position = "none",
        axis.ticks = element_line(colour = "black"),
        plot.title = element_text(hjust = 0.5, size=11),
        plot.subtitle = element_text(hjust = 0.5, size=9, face="italic"),
        panel.background = element_rect(fill = NA),
        panel.grid.major.x = element_line(colour = NA),
        panel.grid.minor.x = element_line(colour = NA),
        axis.text = element_text(colour="black",
                                 face="bold")) +
  ggtitle(label = "Barchart representing number of lnc-RNA pre- and post-capture transcripts\npresent in different numbers of samples (sampleN).")
```

# mT analysis with sample data
## Read data from tsv file
```{r}
# Data DO NOT contain header
read.table(file = '~/Desktop/mT/RNA_mT_Hv3_sample_stats.tsv', sep = '\t', header = FALSE) %>% 
  as_tibble() -> df_rna_samples

colnames(df_rna_samples) <- c('transcript_id',
                              'tissue',
                              'stage',
                              'technology',
                              'capture_status',
                              'biological_replicate',
                              'technical_replicate')
```

## Determine number of transcripts for ea platform (multi - both ont and pacBio)
```{r}
multi_rna_transcripts <- df_rna_samples %>%
  select(transcript_id, technology) %>%
  filter(technology=="ont") %>%
  select(transcript_id) %>%
  intersect(., (df_rna_samples %>%
                  select(transcript_id, technology) %>%
                  filter(technology=="pacBioSII") %>%
                  select(transcript_id)))

technology_rna_df <- data.frame(technology = c("multi", "ont", "pacBioSII"),
                            number_of_transcripts = c(multi_rna_transcripts %>%
                                                        nrow(),
                                                      df_rna_samples %>%
                                                        select(transcript_id, technology) %>%
                                                        filter(technology=="ont") %>%
                                                        select(transcript_id) %>%
                                                        setdiff(., multi_rna_transcripts) %>%
                                                        nrow(),
                                                      df_rna_samples %>%
                                                        select(transcript_id, technology) %>%
                                                        filter(technology=="pacBioSII") %>%
                                                        select(transcript_id) %>%
                                                        setdiff(., multi_rna_transcripts) %>%
                                                        nrow()))

# Plot the results
technology_rna_df %>%
  mutate(percent = prop.table(number_of_transcripts)) %>% 

  ggplot(data=.,
         aes(x=reorder(technology, -percent),
             y=percent,
             fill=technology)) + 
  geom_bar(stat="identity", position = "dodge") +
  geom_text(aes(label = scales::comma(number_of_transcripts)),
            position = position_dodge(width = 1),
            vjust=-0.5,
            size = 3) +
  theme_minimal() +
  ggtitle(label="Number of lnc-RNA transcripts aquired using ont, pacBio or both technologies.") +
  scale_fill_brewer(palette="Set2") +
  ylab('Percentage') +
  xlab('Technology') +
  theme(legend.position = "none") +
  scale_y_continuous(labels = scales::percent_format(),
                     limits = c(0,1)) -> p_rna_technology

#pdf(file = "RNA_transcripts_by_technology.pdf", width = 6, height = 4)
p_rna_technology



##### Stacked barplot
# Plot the results
technology_rna_df %>%
  mutate(percent = prop.table(number_of_transcripts)) %>% 

  ggplot(data=.,
         aes(x=1,#reorder(technology, -percent),
             y=percent,
             fill=factor(technology, levels=c("ont", "multi", "pacBioSII")))) + 
  geom_bar(stat="identity") +
   geom_text(aes(label = paste0(technology, " (", scales::comma(number_of_transcripts), ")")),
             position=position_stack(vjust=0.5),
             size = 4) +
  theme_minimal() +
  scale_fill_manual(values=c("#377EB8", "#4DAF4A", "#FFFF33")) +
  ylab('Percentage') +
  xlab('Technology') +
  theme(legend.position = "none",
        axis.ticks.y = element_line(colour = "black"),
        plot.title = element_text(hjust = 0.5),
        panel.background = element_rect(fill = NA),
        panel.grid.major.x = element_line(colour = NA),
        panel.grid.minor.x = element_line(colour = NA),
        axis.text.x = element_blank(),
        axis.text.y = element_text(colour="black",
                                  face="bold")) +
  ggtitle(label="Number of lnc-RNA transcripts aquired using ont, pacBio or both technologies.") +
  scale_y_continuous(labels = scales::percent_format(),
                     limits = c(0,1)) -> p_rna_technology2

#pdf(file = "RNA_transcripts_by_technology_stacked.pdf", width = 3, height = 4)
p_rna_technology2
```

## Tissues - Correlation matrix
### Pre-capture
```{r}
pre_cap_rna_matrix_df <- df_rna_samples %>%
  filter(capture_status=="pre-capture") %>%    # Only pre-capture
  distinct(transcript_id, tissue, stage) %>%   # Leave only unique tissue per transcript > don't include technology
  mutate(tissue=paste0(tissue," (", stage, ")")) %>% 
  select(transcript_id, tissue) %>% 
  group_by(tissue) %>% 
  summarize(transcripts = list(transcript_id)) %>% 
  cross_join(x=., y=., suffix = c("1", "2")) %>% 
  rowwise() %>% 
  mutate(intersecting = list(intersect(x=transcripts1, y=transcripts2))) %>% 
  mutate(intersecting_no = length(intersecting)) %>% 
  mutate(percentage = intersecting_no/length(transcripts1) * 100)


pre_cap_rna_matrix_df %>% 
  ggplot(data=.) +
  geom_tile(data = pre_cap_rna_matrix_df %>% filter(percentage == 100),
            aes(x=tissue1,
                y=tissue2), fill="#2491AA") +
  geom_tile(data = pre_cap_rna_matrix_df %>% filter(percentage < 100),
            aes(x=tissue1,
                y=tissue2,
                fill = percentage)) +
  geom_text(size=5,
            aes(x=tissue1,
                y=tissue2,
                label = (scales::percent(percentage/100, accuracy = 0.1L)))) +
  scale_fill_gradient(low = "white",
                      high = "#29aac7") +
   theme(axis.title.x = element_blank(),
         axis.title.y = element_blank(),
         axis.ticks = element_line(colour = NA),
         plot.title = element_text(hjust = 0.5),
         panel.background = element_rect(fill = NA),
         panel.grid.major = element_line(colour = NA),
         axis.text.x = element_text(colour="black",
                                    face="bold"),
         axis.text.y = element_text(colour="black",
                                    face="bold")) +
  labs(title = "Lnc-RNA transcripts shared between tissues (pre-capture).", fill = "% of transcripts") -> p_rna_pre_cap_heatmap


#pdf(file="RNA_pre_cap_heatmap.pdf", width = 16, height = 8)
p_rna_pre_cap_heatmap
```

### Post-capture
```{r}
post_cap_rna_matrix_df <- df_rna_samples %>%
  filter(capture_status=="post-capture") %>%    # Only post-capture
  distinct(transcript_id, tissue, stage) %>%   # Leave only unique tissue per transcript > don't include technology
  mutate(tissue=paste0(tissue," (", stage, ")")) %>% 
  select(transcript_id, tissue) %>% 
  group_by(tissue) %>% 
  summarize(transcripts = list(transcript_id)) %>% 
  cross_join(x=., y=., suffix = c("1", "2")) %>% 
  rowwise() %>% 
  mutate(intersecting = list(intersect(x=transcripts1, y=transcripts2))) %>% 
  mutate(intersecting_no = length(intersecting)) %>% 
  mutate(percentage = intersecting_no/length(transcripts1) * 100)

post_cap_rna_matrix_df %>% 
  ggplot(data=.) +
  geom_tile(data = post_cap_rna_matrix_df %>% filter(percentage == 100),
            aes(x=tissue1,
                y=tissue2), fill="#30E48D") +
  geom_tile(data = post_cap_rna_matrix_df %>% filter(percentage < 100),
            aes(x=tissue1,
                y=tissue2,
                fill = percentage)) +
  geom_text(size=5,
            aes(x=tissue1,
                y=tissue2,
                label = (scales::percent(percentage/100, accuracy = 0.1L)))) +
  scale_fill_gradient(low = "white",
                      high = "#2AC77B") +
   theme(axis.title.x = element_blank(),
         axis.title.y = element_blank(),
         axis.ticks = element_line(colour = NA),
         plot.title = element_text(hjust = 0.5),
         panel.background = element_rect(fill = NA),
         panel.grid.major = element_line(colour = NA),
         axis.text.x = element_text(colour="black",
                                    face="bold"),
         axis.text.y = element_text(colour="black",
                                    face="bold")) +
  labs(title = "Lnc-RNA transcripts shared between tissues (post-capture).", fill = "% of transcripts") -> p_rna_post_cap_heatmap


#pdf(file="RNA_post_cap_heatmap.pdf", width = 16, height = 8)
p_rna_post_cap_heatmap

```

### Pre&Post-capture
```{r}
preANDpost_cap_rna_matrix_df <- df_rna_samples %>%
  distinct(transcript_id, tissue, stage) %>%   # Leave only unique tissue per transcript > don't include technology
  mutate(tissue=paste0(tissue," (", stage, ")")) %>% 
  select(transcript_id, tissue) %>% 
  group_by(tissue) %>% 
  summarize(transcripts = list(transcript_id)) %>% 
  cross_join(x=., y=., suffix = c("1", "2")) %>% 
  rowwise() %>% 
  mutate(intersecting = list(intersect(x=transcripts1, y=transcripts2))) %>% 
  mutate(intersecting_no = length(intersecting)) %>% 
  mutate(percentage = intersecting_no/length(transcripts1) * 100)

preANDpost_cap_rna_matrix_df %>% 
  ggplot(data=.) +
  geom_tile(data = preANDpost_cap_rna_matrix_df %>% filter(percentage == 100),
            aes(x=tissue1,
                y=tissue2), fill="#E48430") +
  geom_tile(data = preANDpost_cap_rna_matrix_df %>% filter(percentage < 100),
            aes(x=tissue1,
                y=tissue2,
                fill = percentage)) +
  geom_text(size=5,
            aes(x=tissue1,
                y=tissue2,
                label = (scales::percent(percentage/100, accuracy = 0.1L)))) +
  scale_fill_gradient(low = "white",
                      high = "#C8742A") +
   theme(axis.title.x = element_blank(),
         axis.title.y = element_blank(),
         axis.ticks = element_line(colour = NA),
         plot.title = element_text(hjust = 0.5),
         panel.background = element_rect(fill = NA),
         panel.grid.major = element_line(colour = NA),
         axis.text.x = element_text(colour="black",
                                    face="bold"),
         axis.text.y = element_text(colour="black",
                                    face="bold")) +
  labs(title = "Lnc-RNA transcripts shared between tissues (pre&post-capture).", fill = "% of transcripts") -> p_rna_preANDpost_cap_heatmap


#pdf(file="RNA_preANDpost_cap_heatmap.pdf", width = 16, height = 8)
p_rna_preANDpost_cap_heatmap
```









# JUNK Load data:
```{r}
df <- read.delim("/home/tomasz/Desktop/mT/RNA_mT_Hv3_sample_stats.tsv",
                 sep="\t",
                 header = FALSE)
#View(df)

colnames(df) <- c('tissue',
                  'stage',
                  'technology',
                  'capture_status',
                  'biological_replicate',
                  'technical_replicate')
```

## Create some plots:
```{r}
df %>%
  group_by(technology) %>%
  summarise(total_count=n()) %>% 
  mutate(percent = prop.table(total_count)) %>% 

  ggplot(data=.,
         aes(x=technology,
             y=percent,
             fill=technology)) + 
  geom_bar(stat="identity", position = "dodge") +
  geom_text(aes(label = scales::comma(total_count)), #label = paste0(total_count)),
            position = position_dodge(width = 1),
            vjust=-0.5,
            size = 3) +
  theme_minimal() +
  scale_fill_brewer(palette="Pastel1") +
  ylab('Percentage') +
  xlab('Technology') +
  theme(legend.position = "none") +
  scale_y_continuous(labels = scales::percent_format(),
                     limits = c(0,1)) -> p1

df %>%
  group_by(tissue) %>%
  summarise(total_count=n()) %>% 
  mutate(percent = prop.table(total_count)) %>% 

  ggplot(data=.,
         aes(x=reorder(tissue, -percent),
             y=percent,
             fill=tissue)) + 
  geom_bar(stat="identity", position = "dodge") +
  geom_text(aes(label = scales::comma(total_count)),
            position = position_dodge(width = 1),
            vjust=-0.5,
            size = 2) +
  theme_minimal() +
  scale_fill_brewer(palette="Spectral") +
  xlab('Tissue') +
  ylab('Percentage') +
  theme(legend.position = "none") +
  scale_y_continuous(labels = scales::percent_format(),
                     limits = c(0,0.3)) -> p2

df %>%
  group_by(stage) %>%
  summarise(total_count=n()) %>% 
  mutate(percent = prop.table(total_count)) %>% 

  ggplot(data=.,
         aes(x=stage,
             y=percent,
             fill=stage)) + 
  geom_bar(stat="identity", position = "dodge") +
  geom_text(aes(label = scales::comma(total_count)),
            position = position_dodge(width = 1),
            vjust=-0.5,
            size = 3) +
  theme_minimal() +
  scale_fill_brewer(palette="Set3") +
  ylab('Percentage') +
  xlab('Stage') +
  theme(legend.position = "none") +
  scale_y_continuous(labels = scales::percent_format(),
                     limits = c(0,1)) -> p3

df %>%
  group_by(capture_status) %>%
  summarise(total_count=n()) %>% 
  mutate(percent = prop.table(total_count)) %>% 

  ggplot(data=.,
         aes(x=reorder(capture_status, -percent),
             y=percent,
             fill=capture_status)) + 
  geom_bar(stat="identity", position = "dodge") +
  geom_text(aes(label = scales::comma(total_count)),
            position = position_dodge(width = 1),
            vjust=-0.5,
            size = 3) +
  theme_minimal() +
  scale_fill_brewer(palette="Spectral") +
  xlab('Capture status') +
  ylab('Percentage') +
  theme(legend.position = "none") +
  scale_y_continuous(labels = scales::percent_format(),
                     limits = c(0,1)) -> p4


#svg(bg = "white", width=14, height=6)
#pdf(bg = "white", width=14, height=6)
#pdf(file = "RNA_samples_plot", width = 12, height = 5)
p1 + p2 + p3 + p4
```