---
title: "R Notebook"
output: html_notebook
---


# Read modules
```{r}
library(dplyr)
library(ggplot2)
library(patchwork)
library(purrr)
library(stringr)
```


# Read data from tsv file
```{r}
# Read targets from targetDesign file
# Data contains the header
read.table(file = '~/Desktop/mT/new-masterTable/Hv3_CLS3_targetDesign.tsv', sep = '\t', header = TRUE) %>% 
  as_tibble() -> df_targets
df_targets


# Read targets from masterTable with transcript_ids
# Data contains the header
read.table(file = '~/Desktop/mT/new-masterTable/targets_Hv3_ICmT.tsv', sep = '\t', header = TRUE) %>% 
  as_tibble() -> df_mT
df_mT


# Read targets from masterTable with transcript_ids, target_ids and sample tissue
# Data contains the header
read.table(file = '~/Desktop/mT/new-masterTable/targets_Hv3_ICmT_with_tissues.tsv', sep = '\t', header = TRUE) %>% 
  as_tibble() -> df_ICmT_with_tissues
df_ICmT_with_tissues


# Read transcripts from concatenated catalogues file
# Data DO NOT contain the header
read.table(file = '~/Desktop/mT/new-masterTable/Catalogues/concat_catalogues.tsv', sep = '\t', header = FALSE) %>% 
  as_tibble() %>% 
  rename(transcript_id=V1,
         catalogue=V2)-> df_catalogues
df_catalogues


```

# Check number of targets for ea catalogue (lncRNA only)
```{r}
#pdf(file = "N_targets_per_catalogue.pdf", width = 8, height = 4)
df_targets %>% 
  filter(target_category=="longNonCoding") %>% 
  group_by(target_set) %>% 
  summarise(count=n()) %>% 
  ggplot(aes(x = reorder(target_set, -count), y = count, fill=target_set)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = scales::comma(count)),
            position = position_dodge(width=0.9),
            vjust=-0.25,
            size = 3.5) +
  ylab('Number of targets') +
  ggtitle(label = "Number of lncRNA targets for each catalogue") +
  theme_minimal() +
  #scale_fill_brewer(palette="Paired") +
  theme(axis.title.x=element_blank(),
        axis.text = element_text(colour = "black", size = 11),
        axis.ticks = element_line(colour = "black"),
        panel.background = element_rect(fill = "white", colour = "grey50"),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        legend.position = "none",
        panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5, size=11)
        ) -> p_targeted_catalogues

p_targeted_catalogues
```
# Number of (lncRNA) targets detected in masterTable - DETECTED/UNDETECTED
```{r}
df_targets %>% 
  filter(target_category=="longNonCoding") %>% 
  select(target_id, target_set) %>% 
  left_join(x=.,
            y=(df_mT %>%
                 distinct(target) %>% 
                 mutate(status="detected")),
            by=join_by(target_id==target)) %>% 
  replace(is.na(.), "undetected") %>%
  summarise(count=n(), .by=c(target_set, status)) %>% 
  ggplot(data=.,aes(x=target_set,
                    y=count,
                    fill=status)) +
  geom_bar(stat="identity", position=position_dodge())
```

# Number of (lncRNA) targets detected in masterTable - TOTAL/DETECTED
```{r}
#pdf(file = "N_IC_detected_targets.pdf", width = 8, height = 4)
df_targets %>% 
  filter(target_category=="longNonCoding") %>% 
  select(target_id, target_set) %>% 
  inner_join(x=.,
            y=(df_mT %>%
                 distinct(target) %>% 
                 mutate(status="detected")),
            by=join_by(target_id==target)) %>%
  summarise(count=n(), .by=c(target_set, status)) %>% 
  rbind(df_targets %>% 
          filter(target_category=="longNonCoding") %>% 
          select(target_id, target_set) %>%
          summarise(count=n(), .by=c(target_set)) %>%
          mutate(status="targeted")) %>%
  rbind(df_catalogues %>%
          summarise(count=n(), .by=c("catalogue")) %>% 
          rename(target_set=catalogue) %>% 
          mutate(status="all")) %>% 
  ggplot(data=.,aes(x=reorder(target_set, -count),
                    y=count,
                    fill=reorder(status, -count))) +
  geom_bar(stat="identity", position=position_dodge()) +
  geom_text(aes(label = paste0(round((count/1000), digits=0), " K")),     #scales::comma(count)),
          position = position_dodge(width=0.9),
          vjust=-0.25,
          size = 2.9) +
  theme_minimal() +
  ylab('Number of targets') +
  ggtitle(label = "Number of lncRNA targets for each catalogue") +
  scale_fill_brewer(palette="Blues", direction = -1) +
  theme(axis.title.x=element_blank(),
        axis.text = element_text(colour = "black", size = 11),
        axis.ticks = element_line(colour = "black"),
        panel.background = element_rect(fill = "white", colour = "grey50"),
        legend.title = element_blank(),
        legend.position = c(0.9, 0.9),
        panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5, size=11)
        ) +
  scale_y_continuous()
  #scale_y_continuous(labels = scales::label_number(suffix = " K", scale = 1e-3))
```

In case of Intron Chain Master Table (ICmT) I observed smaller detection rate - fewer target_ids were present in ICmT compared to
default mT. This phenomenon is perfectly normal, because means that some of the targeted transcripts were mono exonic, thus were not
included in ICmT.


# Targeted/detected chart
```{r}
#pdf(file = "N_IC_detected_targets_percent.pdf", width = 8, height = 4)
df_targets %>% 
  filter(target_category=="longNonCoding") %>% 
  select(target_id, target_set) %>% 
  inner_join(x=.,
            y=(df_mT %>%
                 distinct(target) %>% 
                 mutate(status="detected")),
            by=join_by(target_id==target)) %>%
  summarise(count=n(), .by=c(target_set, status)) %>% 
  rbind(df_targets %>% 
          filter(target_category=="longNonCoding") %>% 
          select(target_id, target_set) %>%
          summarise(count=n(), .by=c(target_set)) %>%
          mutate(status="targeted")) %>%
  group_by(target_set) %>% 
  mutate(percentage=count/max(count)) %>% 
  ggplot(data=.,aes(x=reorder(target_set, -count),
                    y=count,
                    fill=reorder(status, -count))) +
  geom_bar(stat="identity", position=position_dodge()) +
  geom_text(aes(label = paste0(scales::comma(count), "\n(",  scales::percent(percentage), ")")),
          position = position_dodge(width=0.9),
          vjust=-0.25,
          size = 2.9) +
  theme_minimal() +
  ylab('Number of targets') +
  ggtitle(label = "Number of lncRNA targets for each catalogue") +
  scale_fill_brewer(palette="Blues", direction = -1) +
  theme(axis.title.x=element_blank(),
        axis.text = element_text(colour = "black", size = 11),
        axis.ticks = element_line(colour = "black"),
        panel.background = element_rect(fill = "white", colour = "grey50"),
        legend.title = element_blank(),
        legend.position = c(0.9, 0.9),
        panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5, size=11)
        ) +
  scale_y_continuous(limits = c(0, 150000), labels=scales::comma)
```

# Targeted/detected chart - ale z naciskiem na %
```{r}
#pdf(file = "N_IC_detected_targets_percent2.pdf", width = 8, height = 4)
df_targets %>% 
  filter(target_category=="longNonCoding") %>% 
  select(target_id, target_set) %>% 
  inner_join(x=.,
            y=(df_mT %>%
                 distinct(target) %>% 
                 mutate(status="detected")),
            by=join_by(target_id==target)) %>%
  summarise(count=n(), .by=c(target_set, status)) %>% 
  rbind(df_targets %>% 
          filter(target_category=="longNonCoding") %>% 
          select(target_id, target_set) %>%
          summarise(count=n(), .by=c(target_set)) %>%
          mutate(status="targeted")) %>%
  group_by(target_set) %>% 
  mutate(percentage=count/max(count)) %>% 
  ggplot(data=.,aes(x=reorder(target_set, -percentage),
                    y=percentage,
                    fill=reorder(status, -count))) +
  geom_bar(stat="identity", position=position_dodge()) +
  geom_text(aes(label = paste0(scales::percent(percentage), "\n(", scales::comma(count), ")")),
          position = position_dodge(width=0.9),
          vjust=-0.25,
          size = 2.9) +
  theme_minimal() +
  ylab('Percentage') +
  ggtitle(label = "Targeted to detected ratio for each catalogue") +
  scale_fill_brewer(palette="Blues", direction = -1) +
  theme(axis.title.x=element_blank(),
        axis.text = element_text(colour = "black", size = 11),
        axis.ticks = element_line(colour = "black"),
        panel.background = element_rect(fill = "white", colour = "grey50"),
        legend.title = element_blank(),
        legend.position = "top",  #c(0.9, 0.9),
        panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5, size=11)
        ) +
  scale_y_continuous(breaks = c(0, 0.25, 0.50, 0.75, 1.00), limits = c(0, 1.11), labels=scales::percent)
```



# Targeted/detected chart - ale z naciskiem na % i podziałem na tkanki
```{r}
#pdf(file = "N_IC_detected_targets_percent2_tkanka.pdf", width = 8, height = 4)
df_targets %>% 
  filter(target_category=="longNonCoding") %>% 
  select(target_id, target_set) %>% 
  inner_join(x=.,
            y=(df_ICmT_with_tissues %>%
                 filter(tissue=="Brain") %>% # Testis, CpoolA, Brain, Placenta, WBlood, Heart, TpoolA, Liver, iPSC
                 filter(technology=="ont") %>% # ont, pacBioSII
                 filter(stage=="Adult") %>% # Adult, Placenta, Embryo
                 filter(capture_status=="post-capture") %>%
                 filter(bioReplicate==3) %>%
                 filter(techReplicate==1) %>%
                 #filter(grepl('chr', chr)) %>%
                 distinct(target) %>%  
                 mutate(status="detected")),
            by=join_by(target_id==target)) %>%
  summarise(count=n(), .by=c(target_set, status)) %>% 
  rbind(df_targets %>% 
          filter(target_category=="longNonCoding") %>% 
          select(target_id, target_set) %>%
          summarise(count=n(), .by=c(target_set)) %>%
          mutate(status="targeted")) %>%
  group_by(target_set) %>% 
  mutate(percentage=count/max(count)) %>% 
  ggplot(data=.,aes(x=reorder(target_set, -percentage),
                    y=percentage,
                    fill=reorder(status, -count))) +
  geom_bar(stat="identity", position=position_dodge()) +
  geom_text(aes(label = paste0(scales::percent(percentage), "\n(", scales::comma(count), ")")),
          position = position_dodge(width=0.9),
          vjust=-0.25,
          size = 2.9) +
  theme_minimal() +
  ylab('Percentage') +
  ggtitle(label = "Targeted to detected ratio for each catalogue (WBlood)") + # Testis, CpoolA, Brain, Placenta, WBlood, Heart, TpoolA, Liver, iPSC
  scale_fill_brewer(palette="Blues", direction = -1) +
  theme(axis.title.x=element_blank(),
        axis.text = element_text(colour = "black", size = 11),
        axis.ticks = element_line(colour = "black"),
        panel.background = element_rect(fill = "white", colour = "grey50"),
        legend.title = element_blank(),
        legend.position = "top",  #c(0.9, 0.9),
        panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5, size=11)
        ) +
  scale_y_continuous(breaks = c(0, 0.25, 0.50, 0.75, 1.00), limits = c(0, 1.11), labels=scales::percent)
```













# Junk below
## Fajny wykres, ale nie o to chodziło
```{r}
#pdf(file = "detected_targets.pdf", width = 8, height = 4)
df_targets %>% 
  filter(target_category=="longNonCoding") %>% 
  select(target_id, target_set) %>% 
  inner_join(x=.,
            y=(df_mT %>%
                 distinct(target) %>% 
                 mutate(status="detected")),
            by=join_by(target_id==target)) %>%
  select(target_set, status) %>% 
  #summarise(count=n(), .by=c(target_set, status)) %>% 
  rbind(df_targets %>% 
          filter(target_category=="longNonCoding") %>% 
          mutate(status="targeted") %>%
          select(target_set, status)) %>%
          #summarise(count=n(), .by=c(target_set)) %>%
  ggplot(data=.,aes(x=as.factor(target_set),
                    fill=as.factor(status))) +
  geom_bar(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..]), position="dodge") +
  geom_text(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..], label=scales::percent(..count../tapply(..count.., ..x.. ,sum)[..x..])),
          stat = "count",
          position = position_dodge(width=0.9),
          vjust=-0.25,
          size = 2.9) +
  theme_minimal() +
  ylab('Number of targets') +
  ggtitle(label = "Number of lncRNA targets for each catalogue") +
  scale_fill_brewer(palette="Blues", direction = -1) +
  theme(axis.title.x=element_blank(),
        axis.text = element_text(colour = "black", size = 11),
        axis.ticks = element_line(colour = "black"),
        panel.background = element_rect(fill = "white", colour = "grey50"),
        legend.title = element_blank(),
        legend.position = c(0.9, 0.9),
        panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5, size=11)
        ) +
  scale_y_continuous(labels = scales::percent)


# ggplot(mtcars, aes(x=as.factor(cyl), fill=as.factor(gear)))+
#   geom_bar(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..]), position="dodge" ) +
#   geom_text(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..], label=scales::percent(..count../tapply(..count.., ..x.. ,sum)[..x..]) ),
#             stat="count", position=position_dodge(0.9), vjust=-0.5)
  

```





```{r}
df_targets %>% 
  filter(target_category=="longNonCoding") %>%
  select(target_id, target_set) %>% 
  filter(target_set=="gencodeLncRna+")
```






```{r}
df_targets %>% 
  filter(target_category=="longNonCoding") %>% 
  select(target_category, target_set) %>% 
  distinct(target_set)
```

```{r}
df_mT %>%
  distinct(target) %>% 
  nrow()
```



```{r}
df_targets %>% 
  filter(target_category=="longNonCoding") %>% 
  select(target_id, target_set) %>% 
  left_join(x=., y=df_mT, by=join_by(target_id==target)) %>% 
  # group_by(target_set) %>% 
  # summarise(count=n())
  
  df_targets %>% 
  filter(target_category=="longNonCoding") %>% 
  group_by(target_set) %>% 
  summarise(count=n()) %>% 
  rename(catalog = target_set) %>% 
  mutate(type="targeted")
  
```

